% The 12pt option is required by the 2001/02 thesis regulations
% Hacked together from the LaTeX template from the maths department of UoM

%Magic latex symbol finder: http://detexify.kirelabs.org/classify.html

%This is just the standard line that tells the virtual "printer" what document we're making.
% Remove the two side option for single-sided printing
\documentclass[12pt,PhD,twoside]{muthesis}

%These are the citation styles.
\usepackage[backend=bibtex,style=nature,citestyle=nature, doi=false,isbn=false,url=false, maxbibnames=99]{biblatex} %science, nature, alphabetic are common choices.
%FYI

% Citations styles = https://www.sharelatex.com/learn/Biblatex_citation_styles
% Bibliography styles = https://www.sharelatex.com/learn/Biblatex_bibliography_styles

\usepackage{graphicx} %use [!ht] (which means ignore sensible protocols and place here) not [h] (which means place here-ish) for figures otherwise they end up at the end of the chapter.
\usepackage{amsmath}
%\usepackage{siunitx}
\usepackage{csquotes}
\usepackage{longtable}
\usepackage[font=footnotesize,labelfont=bf,justification=justified,format=plain,up,textfont=up]{caption} %\usepackage{tabularx}
\usepackage{ctable}
\usepackage{multirow}
\usepackage{enumerate}
\usepackage[acronym]{glossaries}
\usepackage{ragged2e} % allows handling long code lines and words over multiple lines.
\usepackage{rotating} % allows rotated figures

%Special commands
\newcommand{\specialcell}[2][c]{%
  \begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}
\newcommand{\angstrom}{\textup{\AA}}

\addbibresource{references.bib} % This is the file containing all your bibtex references. All reference managers can export to this format.
\graphicspath{{images/}} %Store all your figures here

% Example abbreviation: {the code used in the text} {the short form} {the long form}. For help on introducing nomenclature, see the sharelatex help page. http://tex.stackexchange.com/questions/86666/how-to-create-both-list-of-abbreviations-and-list-of-nomenclature-using-nomencl

\makeglossaries{}
\newacronym{pm}{PM}{Plasma Membrane}
\newacronym{md}{MD}{Molecular Dynamics}
\newacronym{tm}{TM}{Transmembrane}
\newacronym{tmh}{TMH}{Transmembrane Helix}
\newacronym{tms}{TMS}{Transmembrane Segment}
\newacronym{tmp}{TMP}{Transmembrane Protein}
\newacronym{tmd}{TMD}{Transmembrane Domain}
\newacronym{er}{ER}{Endoplasmic Reticulum}
\newacronym{ta}{TA}{Tail Anchor}
\newacronym{gpi}{GPI}{Glycosylphosphatidylinositol}
\newacronym{popc}{POPC}{Palmitoyloleoylphosphatidylcholine}
\newacronym{pdb}{PDB}{Protein Data Bank}
\newacronym{sp}{SP}{Signal Peptide}
\newacronym{snare}{SNARE}{Soluble N-ethylmaleimide-sensitive factor attachment protein receptor}
\newacronym{ks}{K\--S}{Kolmogorov\--Smirnov}
\newacronym{kw}{K\--W}{Kruskal\--Wallis}
\newacronym{gpcr}{GPCR}{G protein\--coupled receptor}
\newacronym{htl}{HTL}{Holotranslocon}
\newacronym{rna}{RNA}{Ribonucleic Acid}
\newacronym{srp}{SRP}{Signal Recognition Particle}
\newacronym{sr}{SR}{Signal Recognition Particle Receptor}
\newacronym{ap}{AP}{Arrest Peptide}
\newacronym{emc}{EMC}{ER Membrane protein Complex}
\newacronym{em}{EM}{Electron Microscopy}
\newacronym{tom}{TOM}{Translocase of the Outer Membrane}
\newacronym{tim}{TIM}{Translocase of the Inner Membrane}
\newacronym{mom}{MOM}{Mitochondrial Outer Membrane}





\begin{document}
\sloppy

\title{Investigating the Recognition and Interactions of Non-Polar \(\alpha\) Helices in Biology}
\author{James Alexander Baker}
\orcid{orcid.org/0000\--0003\--0874\--2298}
\faculty{Chemistry}
\def\wordcount{22,000}


% Uncomment the line below to suppress the `List of Tables' page (optional)
%\tablespagefalse{}

% Uncomment the line below to suppress the `List of Figures' page (optional)
%\figurespagefalse{}

% Uncomment the line below to use a customised Declaration statement
%\def\declaration{All the work in this thesis has been sourced from Google}


\beforeabstract{} % Don't move the abstract, it doesn't like it. The \beforeabstract command creates the title page, a copyright page (default), the tables of contents, tables and figures.


\prefacesection{Abstract}
Non\---polar helices feature prominently in structural biology.
Transmembrane \(\alpha\) helix containing proteins make up around a quarter of all proteins, represent around 40\% of drug targets, and contain some of the most critical proteins required for life as we know it.
Yet they are fundamentally difficult to study experimentally.
This is in part due to the very features that make them so biologically influential; their non\---polar transmembrane helical regions.

By leveraging large data-sets of transmembrane proteins, this thesis is focused on characterising features of transmembrane \(\alpha\) helices en masse, particularly regarding their topology, membrane\---protein interactions, and intramembrane protein interactions.

In this study, we present statistical evidence demonstrating the ``negative\--outside'' rule in opposition to the ``positive\--inside'' rule.
We also identify stabilising amino acid distributions in anchoring transmembrane helices compared to transmembrane helices with function beyond anchoring.

Tail\--anchored proteins are a group of post\--translationally inserting proteins.
In this thesis we show adaptations of residue distributions through the transmembrane helices of tail\--anchored proteins to different membrane environments within the cell (the mitochondria, endoplasmic reticulum, the Golgi, and the plasma membrane), but could not scrutinise a difference between global populations of tail\--anchored proteins in species (mammals, plants, and fungi).
A handful of these proteins are capable of integrating into the membrane without the need for membrane integration proteins.
Structural modelling of transmembrane helices from PTP1b and cytochrome b5 reveals a 3D amphipathic arrangement of residues.
This structural feature may play a role in their spontaneous membrane insertion.

Finally, we find a conserved pattern of typically hydrophobic transmembrane helices neighbouring marginally hydrophobic helices in some families of transmembrane proteins.
This feature corresponds to transmembrane helices that have the potential to cooperate in order to integrate the more polar, but functionally important, transmembrane helix of the pair into the membrane.


\prefacesection{Lay Abstract} %The lay abstract section is optional.

The survival of each of our cells relies on a cellular barrier (called the membrane) to separate themselves from the surrounding environment.
The membrane works by being chemically very different from both the outside environment and the inside of the cell, which in both cases is mostly water.
The membrane is fatty so repels water.

Proteins are the molecular machinery that forms much of the cell structure and shape as well as carrying out many of the cell's routine tasks.
Around a third of our genome codes for membrane\--embedded proteins.
But because these membrane\--proteins are adapted for a life in the water\--repelling cell membrane, they are very hard to study in laboratories which often rely on methods that hold proteins in water\--based solutions.

In this thesis, we focus particularly on the parts of the protein that are embedded in the water repelling membrane.
We computationally analysed the biochemical make\--up of thousands of proteins from openly available biological databases.

This thesis demonstrates three features of membrane proteins:
\begin{itemize}

  \item the radically different evolutionary story that membrane\--bound regions have compared to other proteins; the sacrifices they make for their stability in order to maintain their function, and their optimisation through evolutionary timescales to adapt to the membrane as best they can.

  \item a distinct sub\--group of membrane\--proteins that have a radically different membrane\--insertion mechanism (tail\--anchored proteins) have adaptations in their membrane regions depending on where they are located in the cell.

  \item some types of membrane proteins may use several membrane elements to ensure the least stable, but functionally important, elements are correctly inserted into the membrane.

\end{itemize}

These results will go on to inform more specific studies about membrane proteins.
These findings will provide insight into the causes of some genetic diseases as well as drug targets in the case of pathogenic infections and cancers.



 %But what is the point, what have we learned.



\afterabstract{} % \afterabstract command to insert the declaration.

%The preface section doesn't like being in a different document for some reason. (I think it's just my compiler, there is no reason why it shouldn't work)
\prefacesection{Acknowledgements}
I would like to thank all members of both the Eisenhaber research group, as well as the Curtis and Warwicker research group for their interesting and engaging discussion throughout my PhD.
In particular I would like to acknowledge the patience, guidance, and exceptional supervision from Dr Jim Warwicker, and Professor Stephen High of the University of Manchester and Dr Frank Eisenhaber, and Dr Birgit Eisenhaber, from the A*STAR in Singapore.
I also express my gratitude towards Dr Wing Cheong Wong and Dr Max Hebditch for working with me on many fiddly problems.
I would also like to thank The University of Manchester and the ARAP programme at the Singapore Bioinformatics Institute for funding the project.

Outside of the office I thank my mother, father, my brother Tim, and Emily for their unwavering support of me in undertaking a PhD based for two years on the otherside of the planet.
Being so far from my family and loved ones was certainly the most challenging part of this process.

%It was my uncle Howard's wisdom and philosophical outlook that swayed me to undertake a PhD.
%May he rest in peace.

% \prefacesection{The Author}

% In 2013, I undertook an internship at Chulalongkorn University under the supervision of Dr. Thanyada Rungrotmongkol. The internship contributed to simulating several compounds that could target a viral protease (\(3C^{pro}\)) of both Cocksackie 16 and Enterovirus 72, of which infect X and kill Y people per year. The results showed that several compounds were more potent inhibitors at a protein level than the currently championed drug candidates.

% I have previously completed a Masters in Biochemistry from the University of Liverpool, graduating in 2014. In the final year of the Masters degree, under the supervision of Dr. Dan Rigden, I modelled domains of a transmembrane protein involved in bacterial competence (ComEC) in \textit{Bacillus subtilis}. Structural and sequential analysis of the models revealed a cryptic OB fold in a domain of unknown function, and a potential competence nuclease function in the $\beta$--lactamase--like domain~\cite{Baker2016}.

% During the first year of this PhD studentship, the author presented a part of the work in Berlin at the FEBS annual congress (Baker, J. A. and Warwicker, J. A Bioinformatic Method to Identify Potential SNARE Proteins. \textit{ 40th FEBS Congress} Late Breaker (2015))

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection*{Preamble} %A side note. the "*" hides the section number.
%In the 1950s and 1960s, the field of biological philosophy was still emerging. David Hull writes on the matter in his 1969 article entitled \textit{``What Philosophy of Biology is not''}:

%\begin{displayquote}
%``Periodically through the history of biology, biologists have tried to do a little philosophy...''\cite{Hull1969}
%\end{displayquote}

%I can't help but wish my thesis title was ``The ins-and-outs of greasy peptides''.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% List of publications not necessary.
% \prefacesection{List of publications}

% \subsection*{Journal Articles} % the * hides the section number

% \subsection*{Posters}
%Baker, J. A. and Warwicker, J. A Bioinformatic Method to Identify Potential SNARE Proteins. \textit{ 40th FEBS Congress} Late Breaker (2015)

%\clearpage
\afterpreface{}

%\pagenumbering{arabic}

%This is where we call the chapters. No need to include ".tex". The chapters will be printed to the document in this order. To create a new chapter, make a new file, and add the name of the file below (without the file extension).
\input{chapter_introduction}
\input{chapter_tmcomposition}
\input{chapter_tailanchor}
\input{chapter_multipassfolding}

\input{chapter_conclusions}

% In order for this to work a special build sequence is needed. http://tex.stackexchange.com/a/46732/42423

%This line prints the bibliography ;)
%One of the references has a "-" in the wrong place and screws up the Bibliography. %the planque reference is messed up. It may need changing with every bib.tex update unless the permanent record is changed.
\clearpage
\printbibliography[title={Bibliography}]

% Uncomment the following THREE lines if you do have an Appendix
%\appendix
%\chapter{}
%.........

\end{document}
